Block bootstrap consistency\\under weak assumptions

### 2017-01-10 Paper.tex
[#32/Ems] Fix minor typo in defining the bootstrap conditional expectations
[#33/lem1] Clean up first component of Lemma 1

First, the relevant probability infrastructure.  Let
$(S, \Omega, \pr)$ be a probability space and define the sequence of
sub-sigma-fields $\Omega_n \subset \Omega$. Assume that each
$X_{nt}$, $V_{nt}$, $M_{nj}$, and $u_{nj}$ is $\Omega_n$-measurable,
with $u_{nj}$ the $\operatorname{uniform}(1,\dots,n)$ random variables
that designate the start period of each bootstrap ``block.'' Also
define the \sigmafield\ generated by the stationary bootstrap's block
lengths alone,
\begin{equation}
  \Ms = \sigma(J_{n}, M_{n1},\dots,M_{nJ_n})
\end{equation}
and the conditional probability $\Pm(\cdot) = \pr[\cdot \mid \Ms]$.
(And define $\Pm^*(\cdot) = \pr^*[\cdot \mid \Ms]$, $\Em(\cdot) =
\E^*(\cdot \mid \Ms)$, etc.) An important property is
that $\Ms$ is \emph{independent} of the $X_{nt}$'s, $V_{nt}$'s, and
$u_{nj}$'s, so we can treat any
$M_{nj}$ and $J_n$ terms as constants within $\Em(\cdot)$ and
$\Pm[\cdot]$ and integrate over the unconditional distributions of the
other random variables. This property is especially important because it allows us
to use maximal inequalities and other moment inequalities for
mixingale arrays with only small modifications to construct almost sure bounds on the
conditional moments $\Em$.

@@

Lemma~\ref{L4} implies that
$\sigma_n^{*2}$ and $\hat\sigma_n^{*2}$ both converge to $\sigma^2$ in
probability. This convergence then implies that
\begin{equation}
  \label{eq:23}
  \pr^{*}\big[\sqrt{n}(\bar X_{n}^{*} - \bar X_n) \leq x\big] \to^p \Phi(x/\sigma)
\end{equation}
and
\begin{equation}
\label{eq:24}
  \pr^{*}\big[\sqrt{n}(\bar X_{n}^{*} - \bar X_n) / \hat\sigma_n^{*}
  \leq x\big] \to^p \Phi(x)
\end{equation}
for any $x$.  These results are sufficient for~\eqref{eq:36}
and~\eqref{eq:2} though an argument attributed to Poly{\`a}
that proceeds as follows.
Let $k$ be a finite integer and define $x_i = \sigma \Phi^{-1}(i/k)$ for $i =
0,\dots,k$ (so $x_0 = -\infty$ and $x_k = +\infty$).
For any $x \in [x_i, x_{i+1}]$,
\begin{align*}
  \pr^*\big[\sqrt{n} (\bar X_n^* - \bar X_n) \leq x] - \Phi(x/\sigma)
  &\leq \pr^*\big[\sqrt{n} (\bar X_n^* - \bar X_n) \leq x_{i+1}] - \Phi(x_i/\sigma) \\
  &= \pr^*\big[\sqrt{n} (\bar X_n^* - \bar X_n) \leq x_{i+1}] - \Phi(x_{i+1}/\sigma) + 1/k
\intertext{and}
  \pr^*\big[\sqrt{n} (\bar X_n^* - \bar X_n) \leq x] - \Phi(x/\sigma)
  &\geq \pr^*\big[\sqrt{n} (\bar X_n^* - \bar X_n) \leq x_i] - \Phi(x_{i+1}/\sigma) \\
  &= \pr^*\big[\sqrt{n} (\bar X_n^* - \bar X_n) \leq x_i] - \Phi(x_i/\sigma) - 1/k
\end{align*}
almost surely. Then
\begin{multline*}
  \sup_{x \in (-\infty, +\infty)} \big| \pr^*\big[\sqrt{n} (\bar X_n^* - \bar X_n) \leq x] - \Phi(x/\sigma) \big| \\
  \leq \sup_{i=0,\dots,k} \big|\pr^{*}\big[\sqrt{n}(\bar X_{n}^{*} - \bar X_n) \leq x_i \big] - \Phi(x_i/\sigma) \big| + 1/k
\end{multline*}
almost surely and~\eqref{eq:23} ensures that
\begin{equation*}
  \sup_{i=0,\dots,k} \big|\pr^{*}\big[\sqrt{n}(\bar X_{n}^{*} - \bar X_n)
  \leq x_i \big] - \Phi(x_i/\sigma) \big| + 1/k \to^p 1/k
\end{equation*}
for any finite $k$. Since $k$ is arbitrary,~\eqref{eq:37} holds. Since
Theorem~\ref{T1}'s assumptions ensure that the original array obeys
the \clt,~\eqref{eq:36} holds \citep[Theorem 2]{Jon:97}.  A similar
argument applies to the asymptotic distribution of
$\sqrt{n}(\bar X_n^* - \bar X_n) / \hat\sigma_n^*$, completing the
proof.\qed }

@@

\begin{proof}[Proof of Lemma~\ref{L1}]
  To prove part 1, observe that, for any increasing positive sequence
  $x_n$ such that $x_n p_n \to \infty$,
  \begin{equation*}
    \pr\Big[\max_{i=1,\dots, \lfloor C n p_n \rfloor} M_{ni} \leq x_n\Big] =
    (1 - (1 - p_n)^{x_n})^{\lfloor C n p_n \rfloor} \to \lim \exp(-C n p_n (1 - p_n)^{x_n})
  \end{equation*}
  and $C n p_n (1 - p_n)^{x_n} \to \lim C n p_n e^{-x_n p_n}$. Now, let
  $x_n = n x$ for any positive number $x$.  Then
  \begin{equation*}
    \pr\Big[\max_{i=1,\dots,\lfloor C n p_n \rfloor} M_{ni}/n \leq x \Big] \to
    \lim \exp(-C n p_n e^{-n p_n x}) =
    \exp(0) = 1.
  \end{equation*}
  Since $x$ is arbitrary, $\max_{i=1,\dots,\lfloor C n p_n \rfloor} M_{ni} / n \to^p 0$.

@@

\begin{lem}\label{L2}
  If $\{A_n\}$ is a sequence of events in $\Omega$ then the following
  are equivalent:
  \begin{align*}
  \pr[A_n] &\to 0, &
  \pr^*[A_n] &\to 0 \text{\ in\ }L_1, &
  &\text{and}&
  \Pm^*[A_n] &\to 0 \text{\ in\ }L_1.
  \end{align*}
\end{lem}
\begin{proof}
  Since $|\pr[A_n]| = \E |\pr^*[A_n]| = \E |\pr^*_{\mathcal{M}}[A_n]|$ these
  conditions are equivalent by definition.
\end{proof}

>>>>

First, the relevant probability infrastructure.  Let
$(S, \Omega, \pr)$ be a probability space and define the sequence of
sub-sigma-fields $\Omega_n \subset \Omega$. Assume that each
$X_{nt}$, $V_{nt}$, $M_{nj}$, and $u_{nj}$ is $\Omega_n$-measurable,
with $u_{nj}$ the $\operatorname{uniform}(1,\dots,n)$ random variables
that designate the start period of each bootstrap ``block.'' Also
define the \sigmafield\ generated by the stationary bootstrap's block
lengths alone,
\begin{equation}
  \Ms = \sigma(J_{n}, M_{n1},\dots,M_{nJ_n})
\end{equation}
and the conditional probability $\Pm(\cdot) = \pr[\cdot \mid \Ms]$.
(And define $\Pm^*(\cdot) = \pr^*[\cdot \mid \Ms]$, $\Em^*(\cdot) =
\E^*(\cdot \mid \Ms)$, etc.) An important property is
that $\Ms$ is \emph{independent} of the $X_{nt}$'s, $V_{nt}$'s, and
$u_{nj}$'s, so we can treat any
$M_{nj}$ and $J_n$ terms as constants within $\Em(\cdot)$ and
$\Pm[\cdot]$ and integrate over the unconditional distributions of the
other random variables.\footnote{%
  A referee suggested the following clarification:
  \begin{itemize}
  \item[$\circ$] $\Em^*(\cdot)$ denotes the expectation over the $u_{ni}$, holding
    the sample and $M_{ni}$ fixed.
  \item[$\circ$] $\Em(\cdot)$ denotes the expectation over the sample and
    $u_{ni}$, holding only the $M_{ni}$ fixed.
  \item[$\circ$] $\E^*(\cdot)$ denotes the expectation over the $u_{ni}$ and
    $M_{ni}$, holding the sample fixed.
  \end{itemize}
  The same distinctions hold for the probabilities as well.} %
This property is especially important because it allows us to use
maximal inequalities and other moment inequalities for mixingale
arrays with only small modifications to construct almost sure bounds
on the conditional moments $\Em$.

@@

Lemma~\ref{L4} implies that
$\sigma_n^{*2}$ and $\hat\sigma_n^{*2}$ both converge to $\sigma^2$ in
probability. This convergence then implies that
\begin{equation}
  \label{eq:23}
  \pr^{*}\big[\sqrt{n}(\bar X_{n}^{*} - \bar X_n) \leq x\big] \to^p \Phi(x/\sigma)
\end{equation}
and
\begin{equation}
\label{eq:24}
  \pr^{*}\big[\sqrt{n}(\bar X_{n}^{*} - \bar X_n) / \hat\sigma_n^{*}
  \leq x\big] \to^p \Phi(x)
\end{equation}
for any $x$.  These results are sufficient for~\eqref{eq:36}
and~\eqref{eq:2} since the limiting distribution is continuous,
a standard argument that proceeds as follows \citep[see][Lemma 2.11, for example]{Vaa:00}.
Let $k$ be a finite integer and define $x_i = \sigma \Phi^{-1}(i/k)$ for $i =
0,\dots,k$ (so $x_0 = -\infty$ and $x_k = +\infty$).
For any $x \in [x_i, x_{i+1}]$,
\begin{align*}
  \pr^*\big[\sqrt{n} (\bar X_n^* - \bar X_n) \leq x] - \Phi(x/\sigma)
  &\leq \pr^*\big[\sqrt{n} (\bar X_n^* - \bar X_n) \leq x_{i+1}] - \Phi(x_i/\sigma) \\
  &= \pr^*\big[\sqrt{n} (\bar X_n^* - \bar X_n) \leq x_{i+1}] - \Phi(x_{i+1}/\sigma) + 1/k
\intertext{and}
  \pr^*\big[\sqrt{n} (\bar X_n^* - \bar X_n) \leq x] - \Phi(x/\sigma)
  &\geq \pr^*\big[\sqrt{n} (\bar X_n^* - \bar X_n) \leq x_i] - \Phi(x_{i+1}/\sigma) \\
  &= \pr^*\big[\sqrt{n} (\bar X_n^* - \bar X_n) \leq x_i] - \Phi(x_i/\sigma) - 1/k
\end{align*}
almost surely. Then
\begin{multline*}
  \sup_{x \in (-\infty, +\infty)} \big| \pr^*\big[\sqrt{n} (\bar X_n^* - \bar X_n) \leq x] - \Phi(x/\sigma) \big| \\
  \leq \sup_{i=0,\dots,k} \big|\pr^{*}\big[\sqrt{n}(\bar X_{n}^{*} - \bar X_n) \leq x_i \big] - \Phi(x_i/\sigma) \big| + 1/k
\end{multline*}
almost surely and~\eqref{eq:23} ensures that
\begin{equation*}
  \sup_{i=0,\dots,k} \big|\pr^{*}\big[\sqrt{n}(\bar X_{n}^{*} - \bar X_n)
  \leq x_i \big] - \Phi(x_i/\sigma) \big| + 1/k \to^p 1/k
\end{equation*}
for any finite $k$. Since $k$ is arbitrary,~\eqref{eq:37} holds. Since
Theorem~\ref{T1}'s assumptions ensure that the original array obeys
the \clt,~\eqref{eq:36} holds \citep[Theorem 2]{Jon:97}.  A similar
argument applies to the asymptotic distribution of
$\sqrt{n}(\bar X_n^* - \bar X_n) / \hat\sigma_n^*$, completing the
proof.\qed }

@@

\begin{proof}[Proof of Lemma~\ref{L1}]
  To prove part 1, take $x$ to be an arbitrary positive number. Then
  \begin{equation*}
    \pr\Big[\max_{i=1,\dots, \lfloor C n p_n \rfloor} M_{ni} \leq n x \Big]
    = (1 - (1 - p_n)^{n x})^{\lfloor C n p_n \rfloor}
    = \exp(-C n p_n e^{-x n p_n}) \times (1 + o(1))
  \end{equation*}
  since $(1 - p_n)^{x/p_n} \to e^{-x}$. In addition, $n p_n e^{-n p_n} \to 0$
  by assumption, so
  \begin{align*}
    \pr\Big[\max_{i=1,\dots, \lfloor C n p_n \rfloor} M_{ni}/n \leq x \Big]
    \to \exp(0) = 1.
  \end{align*}
  Since $x$ is arbitrary, this completes the proof.

@@

\begin{lem}\label{L2}
  If $\{A_n\}$ is a sequence of events in $\Omega$ then the following
  are equivalent:
  \begin{align*}
  \pr[A_n] &\to 0, &
  \pr^*[A_n] &\to 0 \text{\ in\ }L_1, &
  &\text{and}&
  \Pm^*[A_n] &\to 0 \text{\ in\ }L_1.
  \end{align*}
\end{lem}
\begin{proof}
  Since $\pr[A_n] = \E |\pr^*[A_n]| = \E |\pr^*_{\mathcal{M}}[A_n]|$ these
  conditions are equivalent by definition.
\end{proof}
